---
output: github_document
---

# olsLM

The **olsLM** package is a minimal and educational implementation of
ordinary least squares (OLS) linear regression in R.  
It illustrates how linear models can be constructed from scratch without relying
on the built-in `lm()` function.

The package provides:

- `olsLM()` — OLS fitting function (formula or matrix interface)  
- `summary()` method — prints regression coefficients, t-tests, R-squared  
- `plot()` method — diagnostic plots (residuals, Q–Q plot, scale-location, leverage)

---

# Installation

```{r}
install.packages("devtools")
devtools::install_github("mkong2203/olsLM", build_vignettes = TRUE)

library(olsLM)
```

---


# Features

##  Linear regression — `olsLM()`

The main function `olsLM()` fits a linear regression model.

## Summary output — `summary(fit)`

The summary method reports:

1. model call

2. coefficient table (Estimate, Std.Error, t-value, p-value)

3. residual standard error

4. R-squared and adjusted R-squared

##  Diagnostic plots — `plot(fit)`
The plot() method produces the four classical regression diagnostic plots:

1. Residuals vs Fitted

2. Normal Q–Q Plot

3. Scale–Location Plot

4. Residuals vs Leverage

---

# Example

### **Formula interface**

```{r}
fit <- olsLM(mpg ~ wt + hp, data = mtcars)
summary(fit)
plot(fit)
```

### **Matrix interface**

```{r}
X <- as.matrix(mtcars[, c("wt", "hp")])
y <- mtcars$mpg
fit <- olsLM(X, y)
summary(fit)
plot(fit)
```

### **Returned values**

The fitted model stores regression outputs, accessible via $:

```{r}
fit$coefficients    # regression coefficients
fit$fitted          # fitted values
fit$residuals       # ordinary residuals
fit$leverage        # hat values
fit$r_standard      # standardized residuals
fit$r_student       # internally studentized residuals
fit$r_student_ext   # externally studentized residuals
fit$r_squared       # R-squared
fit$adjusted_r_squared # adjusted R-squared
```


